---
title: "Analyzing World Dev. Indicators for Top 24 HDI countries & Bitcons' Predictions"
author: "Julia Chałasiak"
date: "`r format(Sys.time(), '%d/%m/%y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Extended Summary
On of the most known measures for ranking countries is HDI - Human Development Index.  It takes into account: life expectancy at birth (in years), expected years of schooling, mean years of schooling and gross national income per capita. In this report we decided to choose top 24 countries from Latest Human Development Index Ranking and take a closer look at additional information from World Development Indicators in their context. Moreover, we tried to find correlations between Bitcoins' data from 2009 to 2021, S.P.Composite dataset and some of the mentioned countries. Finally, short-term prediction for Bitcoin Market Price USD was made. 

```{r message=FALSE, warning=FALSE, include=FALSE}
libraries <- c("dplyr", "tidyr", "ggplot2", "readxl", "imputeTS", "vtable", "skimr", "DT", "caret", "gganimate", "ggpubr") 

r = getOption("repos")
r["CRAN"] = "http://cran.us.r-project.org"
options(repos = r)

for (l in libraries) {
  if (!require(l)) install.packages(l); library(l, character.only = TRUE)
}

```

### HDI Top
Top 24 countries according to http://hdr.undp.org/en/content/latest-human-development-index-ranking. Because Liechtenstein had many missing values in WDI dataset, we decided to change it to the next top country in the index - France. 
```{r}
hdiTop <- c('Norway', 'Ireland', 'Switzerland', 'Iceland', 'Hong Kong', 'Germany', 'Sweden', 'Netherlands', 'Australia', 'Denmark', 'Finland', 'Singapore', 'United Kingdom', 'New Zealand', 'Belgium', 'Canada', 'United States', 'Austria', 'Israel', 'Japan', 'Slovenia', 'Luxembourg', 'Korea, Rep.', 'Spain', 'France')
```


### Libraries
During generation of this report we used the following libraries:
```{r}
libraries
```
## Datasets Reading

```{r datasets_reading, message=FALSE, warning=FALSE}
currExchRates <- read.csv("./datasets/CurrencyExchangeRates.csv")
goldPrices <- read.csv("./datasets/Gold prices.csv")
spComposite <- read.csv("./datasets/S&P Composite.csv")
worldDevInd <- read_excel("./datasets/World_Development_Indicators.xlsx", trim_ws=TRUE, na = "..")
worldDevInd <- data.frame(worldDevInd)
btcMetadata <- read.csv("./datasets/BCHAIN_metadata.csv")
btcDiff <- read.csv("./datasets/BCHAIN-DIFF.csv")
btcHrate <- read.csv("./datasets/BCHAIN-HRATE.csv")
btcMkpru <- read.csv("./datasets/BCHAIN-MKPRU.csv")
btcTrvou <- read.csv("./datasets/BCHAIN-TRVOU.csv")
```

## Datasets Preparation

### Currency Exchange Rates
In order to clean the Currency Exchange Rates Dataset we inserted 0's in Euro currency before '98 and used special function to fill empty individual places by making mean from adjacent values.

```{r tidy_currency}
CurrExchRatesClean <- currExchRates %>% select(which(colMeans(is.na(.)) < 0.2))
CurrExchRatesClean <- CurrExchRatesClean %>% mutate(Euro = ifelse(Date < as.Date("1998-10-30", format="%Y-%m-%d"), 0, Euro))
CurrExchRatesClean <- CurrExchRatesClean %>% mutate(Date = as.Date(Date, format= "%Y-%m-%d"))
CurrExchRatesClean <- na_ma(CurrExchRatesClean, k = 1)

st(CurrExchRatesClean)
```

### Gold Prices

```{r tidy_gold}
goldPricesClean <- goldPrices %>% mutate(
  EURO..AM. = ifelse(Date < as.Date("1998-12-31", format="%Y-%m-%d"), 0, EURO..AM.),
  EURO..PM. = ifelse(Date < as.Date("1998-12-31", format="%Y-%m-%d"), 0, EURO..PM.),
  USD..PM. = ifelse(Date < as.Date("1968-03-14", format="%Y-%m-%d"), 0, USD..PM.),
  GBP..PM. = ifelse(Date < as.Date("1968-03-14", format="%Y-%m-%d"), 0, GBP..PM.))

goldPricesClean <- goldPricesClean %>% mutate(Date = as.Date(Date, format= "%Y-%m-%d"))
goldPricesClean <- na_ma(goldPricesClean, k = 1)

st(goldPricesClean)
```

### World Development Indicators and Top Chosen Countries
World Development Indicators' Dataset required more transformations than the rest. After creating the Years column top HDI countries for our analysis were filtered. We selected only these indicators which had less than 10% of NA's for filtered countries. The rest of missing values were filled with 0's.

```{r tidy_index, warning=FALSE}

worldDevInd$Series.Name<-gsub("\\$","D",worldDevInd$Series.Name)

worldDevIndClean <- worldDevInd %>% select(-Series.Code) %>% 
  pivot_longer('X1970..YR1970.':'X2020..YR2020.', names_to = "Year", values_to = "Value", values_drop_na = FALSE) %>%
  pivot_wider(names_from = "Series.Name", values_from = Value) %>% unchop(everything())


worldDevIndClean <- worldDevIndClean %>% 
  mutate(Year = substr(Year, 2, 5)) %>% 
  filter(grepl('', Country.Code)) %>%
  relocate(Year, .after = Country.Code)


trial <- as.data.frame(worldDevIndClean)
trial <- worldDevIndClean %>% filter(Country.Name %in% hdiTop)

trial <- trial %>% select(which(colMeans(is.na(.)) < 0.10))


trial[is.na(trial)] <- 0
wdiTop <- trial

datatable(skim(wdiTop %>% dplyr::select(-Country.Name, -Country.Code, -Year)) %>% dplyr::select(skim_variable, numeric.mean, numeric.sd, numeric.p0, numeric.p25, numeric.p75, numeric.p100) %>% dplyr::rename(Variable=skim_variable, Mean=numeric.mean, "Std. Dev."=numeric.sd, Min=numeric.p0, "Pctl. 25"=numeric.p25, "Pctl. 75"=numeric.p75, Max=numeric.p100))
```

### S.P. Composite
```{r tidy_sp_composite}
spCompositeClean <- spComposite %>% 
  mutate(Cyclically.Adjusted.PE.Ratio = ifelse(Year < as.Date("1880-12-31", format="%Y-%m-%d"), 0, Cyclically.Adjusted.PE.Ratio)) %>% rename(Date = Year)
spCompositeClean <- spCompositeClean%>% mutate(Date = as.Date(Date, format= "%Y-%m-%d"))
spCompositeClean <- na_ma(spCompositeClean, k = 1)

st(spComposite)
```

### Bitcoins' Data
We read four datasets containing information about Bitcoins: 
Trvou - Bitcoin USD Exchange Trade Volume,
Mkpru - Bitcoin Market Price USD,
Diff - Bitcoin Difficulty,
Hrate - Bitcoin Hash Rate,
and decided to gather them into one large set:

```{r tidy_bitcoin, echo=TRUE, message=FALSE, warning=FALSE}
btcTrvou <- btcTrvou %>% rename(Trvou = Value)
btcMkpru <- btcMkpru %>% rename(Mkpru = Value)
btcDiff <- btcDiff %>% rename(Diff = Value)
btcHrate <- btcHrate %>% rename(Hrate = Value)
btcAllClean <- btcTrvou %>% left_join(btcMkpru) %>% left_join(btcHrate) %>% left_join(btcDiff)
btcAllClean <- btcAllClean %>% mutate(Date = as.Date(Date, format= "%Y-%m-%d"))
btcAllClean <- na_ma(btcAllClean, k = 1)

st(btcAllClean)
```


## Attributes analysis
In this section we focused mainly on indicators for chosen countries. Later we presented overviews for Bitcoins, Gold Prices and animations for S.P. Composite.
```{r}

```


### World Development Index - Top Chosen based on HDI
We focused on indicators from 2019 because the HDI rank included HDI values from 2019.

```{r attr_analysis_wdi, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}

chosenIndicators <- wdiTop %>% dplyr::select(-`Total greenhouse gas emissions (kt of CO2 equivalent)`,
                                    -`Survival to age 65, female (% of cohort)`,
                                    -`Survival to age 65, male (% of cohort)`,
                                    -`Primary school starting age (years)`,
                                    -`Population, male (% of total population)`,
                                    -`Population, female (% of total population)`,
                                    -`Population, female`,
                                    -`Population, male`,
                                    -`Nitrous oxide emissions (thousand metric tons of CO2 equivalent)`,
                                    -`Methane emissions (kt of CO2 equivalent)`,
                                    -`Methane emissions in energy sector (thousand metric tons of CO2 equivalent)`,
                                    -`CO2 emissions (metric tons per capita)`,
                                    -`CO2 emissions (kt)`
                                    
                                    )  %>% filter(Year == 2019) 

#parametersWdiTop <- c("Urban population growth (annual %)")
colNames <- names(chosenIndicators)[4:40]
for(i in colNames){
  print(ggplot(data = chosenIndicators, aes(x = Country.Name, y =.data[[i]], fill=as.factor(Country.Name))) +
    geom_bar(stat = "identity") + coord_flip() + theme_minimal() + labs(fill='Country', x='Country'))
}


```

### Bitcoins
```{r bitcoin_graphs}

btcAnalysis <- btcAllClean %>% pivot_longer(cols='Trvou':'Diff', names_to = 'Param', values_to = 'Value')
ggplot(btcAnalysis, aes(x=Date, y=Value, color = as.factor(Param))) + geom_line() + facet_wrap(~Param, scales = "free_y") + theme_minimal() + labs(colour = "Bitcoin's Parameter", title = "Bitcoin's Overview")

```

### Gold Prices
```{r attr_analysis_gold}

GoldAnalysis <- goldPricesClean %>% pivot_longer(cols = 'USD..AM.':'EURO..PM.', names_to = 'Curr', values_to = 'Value')
ggplot(data = GoldAnalysis, aes(x = Date, y = Value, color=as.factor(Curr))) +
    geom_point(size=0.1) + facet_wrap(~Curr) + theme_minimal() + labs(color='Currency', title='Gold Prices\' Overview')

```

### S.P. Composite

```{r attr_analysis_spcomposite, echo=FALSE, message=FALSE, warning=FALSE}
spAnalysis <- spCompositeClean %>% pivot_longer(cols='S.P.Composite':'Cyclically.Adjusted.PE.Ratio', values_to = 'Value', names_to = 'Param')

earn <- spAnalysis %>% filter(Param %in% c('Earnings', 'Real.Earnings'))
ggplot(data = earn, aes(x = Date, y = Value, color=as.factor(Param))) +
    geom_line() + facet_wrap(~Param) + labs(color='Type of Earnings') + theme_minimal() + transition_reveal(Date)


```
```{r echo=FALSE, message=FALSE, warning=FALSE}
earn <- spAnalysis %>% filter(Param %in% c('Dividend', 'Real.Dividend'))
ggplot(data = earn, aes(x = Date, y = Value, color=as.factor(Param))) +
    geom_line() + facet_wrap(~Param) + labs(color='Type of Dividend') + theme_minimal() + transition_reveal(Date)
```
```{r}
ggplot(data = spCompositeClean, aes(x = Date, y = S.P.Composite)) +
    geom_line(color='blue') + theme_minimal() + labs(title='S.P.Composite through Years') + transition_reveal(Date)
```

## Correlations
In this section we tried to find correlations using different datasets in order to make better predictions for Bitcoins later in the report. Firstly, some attempts were made with only Bitcoins' data. Then we used S.P. Composite but this significantly limited our possible training and testing sets thus we decided to not use this in predictions. Finally, we looked for correlations between bitcoins and some indicators for top HDI countries. What is important - Bitcoins' dataset was based on days and WDI dataset was based on years. To compare values we calculated means for Bitcoins' years. However, no important correlation was found there.
```{r}

```


### Correlation between Hrate, Diff, Trvou and Mkpru
```{r corr_btc}
btcPlusSP <- btcAllClean %>% inner_join(spCompositeClean, on='Date')
btcPlusGold <- btcAllClean %>% inner_join(goldPricesClean, on='Date')


ggscatter(btcAllClean, x = "Mkpru", y = "Hrate", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Mkpru", ylab = "Hrate", color = 'plum4')

ggscatter(btcAllClean, x = "Mkpru", y = "Diff", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Mkpru", ylab = "Diff", color = 'plum3')

ggscatter(btcAllClean, x = "Mkpru", y = "Trvou", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Mkpru", ylab = "Trvou", color = 'plum2')

```



### Correlations for Bitcoins' Data Including S.P. Composite
```{r correlations}


ggscatter(btcPlusSP, x = "Mkpru", y = "S.P.Composite", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Mkpru", ylab = "S.P.Composite", color='maroon2')

ggscatter(btcPlusSP, x = "Mkpru", y = "CPI", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Mkpru", ylab = "CPI", color='indianred3')

ggscatter(btcPlusSP, x = "Mkpru", y = "Earnings", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Mkpru", ylab = "Earnings", color='coral2')

```

### Correlations for Bitcoins' Data Including Arbitrally Chosen Indicators from WDI
```{r corr_wdi}


btcYearsSummary <- btcAllClean %>% mutate(Year = format(as.Date(Date, format="%d/%m/%Y"),"%Y")) %>% select(-Date) %>% group_by(Year) %>% dplyr::summarize(MeanDiff = mean(Diff), MeanHrate=mean(Hrate), MeanMkpru=mean(Mkpru), MeanTrvou=mean(Trvou)) %>% filter(Year > 2008 & Year <2021)

btcPlusCountry <- btcYearsSummary %>% inner_join(wdiTop, on='Year') 

for (country in hdiTop) {
  countryExample <- btcPlusCountry %>% filter(Country.Name == country)
  check <- cor(x = countryExample$MeanMkpru, y = countryExample$`GDP per capita (current USD)`, method = "pearson", use = "everything")
  if (check < 0.7 | is.na(check)) {next}
  print(ggscatter(countryExample, x = "MeanMkpru", y = "GDP per capita (current USD)", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "MeanMkpru", ylab = "GDP per capita (current USD)", color='orchid2', title = country))
}

```
```{r corr_wdi2}
for (country in hdiTop) {
  countryExample <- btcPlusCountry %>% filter(Country.Name == country)
  check <- cor(x = countryExample$MeanMkpru, y = countryExample$`Urban population growth (annual %)`, method = "pearson", use = "everything")
  if (check < 0.7 | is.na(check)) {next}
  print(ggscatter(countryExample, x = "MeanMkpru", y = "Urban population growth (annual %)", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "MeanMkpru", ylab = "Urban population growth (annual %)", color='maroon1', title = country))
}
```


## Bitcoins' Short-Term Prediction

```{r model, warning=FALSE}

set.seed(23)

btcAllClean2 <- btcAllClean %>% select(-Date)

inTraining <- createDataPartition(y = btcAllClean2$Mkpru, p = .70, list = FALSE)
train <- btcAllClean2[inTraining,]
valid_and_test  <- btcAllClean2[-inTraining,]

inTesting <- createDataPartition(y = valid_and_test$Mkpru, p = .50, list = FALSE)
valid <- valid_and_test[ -inTesting,]
test <- valid_and_test[ inTesting,]


ctrl <- trainControl(
    method = "repeatedcv",
    number = 2,
    repeats = 5)


fit <- train(Mkpru ~ .,
             data = train,
             method = "rf",
             trControl = ctrl,
             ntree = 10)

fit

pred <- predict(fit, newdata = test)
preddf <- data.frame(Pred = pred, Mkpru = test$Mkpru)

datatable(preddf)

print(ggscatter(preddf, x = "Mkpru", y = "Pred",
           
           xlab = "Mkpru", ylab = "Predicted Mkpru", color='maroon1') + geom_smooth())


```






